---
title: "STAT545 Assignment B1"
output: github_document
---

## Writing, Using, and Testing Functions


```{r}
#Load these libraries to run the code contained in this document

library(tidyverse)
library(dplyr)
library(palmerpenguins)
library(datateachr)
library(testthat)
```

### Writing a Function 


While exploring the *vancouver_trees* dataset from the **datateachr** package to complete MDA 1&2 for STAT545A, I found myself repeatedly using several code chunks. One of these was a group_by() %>% summarise() workflow that I used to calculate the number of observations for a given group or groups of variables within the dataset. I decided to write a function that would bundle this workflow and include the option of filtering by a specified condition, so that the function will have multiple uses - counting the number of observations based on one or more grouping variables, or the number of observations that meet a specified condition, or the number of observations based on one or more grouping variables that meet a specified condition.

```{r}

#Function count_by_groups_filter_by_condition()

count_by_groups_filter_by_condition <- function(df, groups, cond) { 
  df %>% 
    group_by(pick({{ groups }})) %>% 
    filter({{ cond }}) %>% 
    summarise(n=n())
}

#' Count observations by groups and/or those which meet a condition
#' 
#' This is a function that summarises the number of observations for one or more variables within a data frame. It can be used to compute the number of  observations of one or more grouped variables, the number of observations that meet a specified condition, or a combination of the two.
#'
#'@param df The first argument is named 'df' as a shorthand for data frame. This is a required argument.
#'
#'@param groups This argument is named 'groups' to refer to the grouping variable(s) that will be input to the group_by function. It must be an object or use the form c(x,y) if multiple variables are used for grouping. It is not required.
#'
#'@param cond This argument is named 'cond' as a shorthand for condition, to refer to the condition that will be specified in the filter function. This argument must be a logical vector (eg sex=="female"). It is not required.
#'
#'@return An object of the same type as `.df`containing the grouped variables and the corresponding number of observations (n) for each row; if no grouping was  applied and only a filtering condition was used, the output will be a 1x1 object containing the number of observations (n) that meet the condition.
#'
#'  The output has the following properties:
#' * Each group is summarised down to one row.
#' * Columns are not modified; column "n" for observations is created.
#' * Data frame attributes are preserved.
#'
#'#'@examples Below are three examples of how to use this function using datasets contained in the **datateachr** and **penguins** packages.

```

### Examples 


It is good practice to check whether a new function is working as intended by running some examples.
```{r}

#Example 1: Vancouver Trees

#In the first example, I will demonstrate how to use the function to group by multiple variables and filter by a condition in the vancouver_trees dataset. I will group by neighbourhood name and height range and specify the condition to be date planted > 2000-01-01 to find the number of trees in each neighbourhood for each height range that were planted after 2000-01-01.

count_by_groups_filter_by_condition(df=vancouver_trees, c(neighbourhood_name, genus_name), cond= date_planted >"2000-01-01")

```

```{r}

#Example 2: Palmer Penguins

#In this example, I will just use the grouping part of the function and I won't include a condition argument to demonstrate that the function can work just by grouping. I will group by island and species and the output should provide the number of observations for each island/species group.

count_by_groups_filter_by_condition(df=penguins, c(island, species))
```

```{r}

#Example 3: Palmer Penguins

#In the last example, I will specify a condition and no groups. I will set the condition to select only female observations, and the output should be a 1x1 dataframe with n = number of female observations.

count_by_groups_filter_by_condition(df=penguins, cond=sex=="female")
```

### Testing the Function 


Next I will write formal tests for my function to ensure that it will give errors when I expect. I will use functions from the **testthat** package. 

```{r}

#Test 1: Does the function give an error when the input is numeric?

test_that("numeric vector causes error", {
  expect_error(count_by_groups_filter_by_condition(15, groups=8))
  })

```

```{r}

#Test 2: Does the function give an error if no data frame is specified?

test_that("no data frame argument causes error",{
  expect_error(count_by_groups_filter_by_condition(groups=species))
  })

```


```{r}
#Test 3: Does the function give an error if the condition is not a logical vector?

test_that("condition as object causes error",{
  expect_error(count_by_groups_filter_by_condition(penguins, c(species, sex), cond=island))
  })

```
